Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-12-09 patchy-review-responses.tex

\vskip 1em

\begin{quote}
  As can you see from the reviews, two general points were raised by all
  reviewers and I feel that thoroughly addressing both of these points poses a
  prerequisite for further consideration of the manuscript:

  (i) There needs to be some sort of numerical verification (e.g. with
  individual-based simulations) to test the main conclusions, check their
  robustness, and identify the parameter ranges for which they hold.
\end{quote}

We agree with this comment, to test the validity of our model we
coded an individual level model of migration and selection. 
We have done a large number of simulations, and compared the results to the theory;
the main comparison occurs in \ref{fig:sim_times} and figures~\ref{fig:sim_probs};
we have also added in supplementary section \ref{apx:parameter_tables} 
visualizations of the simulations at various parameter values
that we think will help some readers in verifying the assumptions.

@@

\begin{quote}
  (ii) The presentation of the main result needs to be improved. In particular,
  all reviewers feel that the main result can be derived in a simpler and more
  intuitive way. I agree with them that at present the manuscript is overly
  technical and would benefit strongly from a more intuitive derivation of the
  results and discussion of the assumptions under which they hold. The reviewers
  provide detailed suggestions on how this can be achieved.
\end{quote}

We have rewritten the method section to simplify the presentation. 
Before writing the paper, we had initially explored a number of simpler ways of developing the
argument (along similar the lines to those suggested by the reviewers), 
and went around in circles for quite some time
before settling on the argument given in the first draft.
As suggested by the third reviewer, 
we have moved the most difficult part of the argument to the Materials \& Methods,
replacing it in the Results with a more \textit{ad hoc} argument
along the lines of that suggested by the reviewers.
We do think the more formal argument is more solid, however,
and therefore necessary to keep; 
as discussed below in our response to the reviewers, various seemingly reasonable arguments
give different answers for subtle reasons.

@@

Other changes of note are:
\begin{itemize}

  \item Several reviewers pointed out that in approximating $\E[1-(1-p_e)^K] \approx \E[K] p_e$
    we were neglecting the case that $s_m \ll s_p$.
    This omission was sloppy on our part,
    which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.

  \item We have also improved our discussion of the limits to the theory,
    mostly in the (short) section \nameref{ss:assumptions}.
    The main point of note was that, indeed as Reviewer 3 pointed out,
    we were running tests in regions where $s_m$ was too small for the theory to apply,
    resulting in $\migrate$ that was not monotonic with $s_m$.
    Happily, the natural restriction of $R \sqrt{s_m}/\sigma > 1$ coincides with the condition
    for $\migrate$ to decrease with $s_m$;
    we don't have good intuition for why the formula should increase in the region where it is inapplicable,
    but this seems beside the point.
\gc{Graham is not sure what you mean by ``we were running tests in regions'' }

@@

%%%%%%%%
\point{long range migration}{All results depend exponentially on the distance, which
      is a consequence of the model only allowing for short range dispersal. Rare
    long range dispersal will alter all these rates dramatically, see for example
    recent work by Hallatschek and Fisher.
}

\reply
This was already discussed in the discussion, but we've added a note about it earlier on, in particular adding a citation to the nice paper of Hallatschek \& Fisher.
\llname{rr:longtailed}

%%%%%%%%
\point{Fig 4:}{ Why not use natural units for the axis of these plots,
  $R\sqrt(s)/\sigma$, $\rho \mu A/s$.
  }

\reply
Good idea, but we've decided to cut this figure in the resubmission.

@@

%%%%%%%%
\point{in the discussion on page 10 "probability of parallel adaptation.."}{
    It should be noted that many of the scenarios at high mutation influx are quite
    hypothetical. While the colonization of migrants might dominate over local
    adaptation, this argument assumes that the migrants are at equilibrium
    frequency. The transient to reach this equilibrium might be much longer than
    the time required for de novo local adaptation.
}

\reply
We don't know of even a poor approximation to the time to reach migration-selection equilibrium,
unfortunately. However, we now pointed out in the section ``\nameref{ss:assumptions}'' that we're missing this point,
and discuss there how the time lag to reach this equilbrium contributes to discrepancies between theory and simulations.

\gc{seems to graham that we should spread to migration-selection
  equilibrium as the fisher-kpp. Suggests that patch width$/sqrt{s}
  \sigma$ wouldn't be a bad guess.}

@@

%%%%%%%%
\point{}{ 
    And speaking of symbols, I think it’s a bit nicer to write things in terms
    of the characteristic patch length $a \sim A^{1/d}$, rather than the area. This
    avoids having to talk about one-dimensional areas \ldots
}

\reply
Good point; we already have $w$ defined for this purpose;
and $A$ appears naturally in $\mutrate$,
but we've tried to make things more intuitive.

%%%%%%%%
\point{}{ 
     It would be nice to have a clear, concise summary of the quantitative
     assumptions needed for the results (all the stuff like $a \gg \sigma/\sqrt{2s_m}$ or
     whatever).
 }

\reply
See text added in Section ``\nameref{ss:assumptions}''.

@@

\reply
This is a very interesting point,
and the basic tool of a diffusion approximation seems to be the same in both.
However, something is not matching up,
because \citet{barton1979geneflow} finds that the length of the region of differentiated genome 
decays with $1/\sqrt{T}$,
whereas with our approximation the length of the differentiated haplotype decays as $1/T$.
The difference may be that Barton considers LD and we consider a haplotype unbroken by recombination.
There are also other assumptions in Barton's paper that it's applicability unclear.
On the balance, and considering the other reviewers' comments,
we've cut this section and discussed it a bit in the discussion. 
\llname{rr:haplotype_discussion}

%%%%%%%%
\point{In “Applications”,}{the authors spend a lot of time introducing the mice
    and calculating the values for Fig. 6, but then they don’t show where the
    actual data lie on those plots. Is there enough information to do some
    kind of comparison? How close are the nearest samples to each other?
    Have no examples of alleles spreading by migration been found? If not,
    how much of a bound does this put on the parameter values? Is there any
    data on the genetic map length of the region of reduced diversity around
    the selected locus?
}

\reply
At the moment it is only known that the same MC1R alleles do not
contribute to color polymorphism on a number of other patches. We know
from personal discussion with Michael Nachman that another color
polymorphism allele, in a different gene, has been mapped on another
patch. We are currently talking to Nachman about a broader-scale
effort to identify other causal alleles motivated in part by the work described
in the paper. However, this work is too preliminary to offer any thing
certain.

 There is no data about the genetic map in this species, but
as a rough rule of thumb 1cM is roughly 1Mb in a number of mammalian
species, we have added this guide to the figure caption \llname{rr:cMMb}

@@

        \item[C.]  The paragraph starting “Also note that. . . ” on p11.
    \end{itemize}
}

\reply
We moved \nameref{ss:simulations} to Materials and Methods,
modified the second sentence,
and rearranged the structure around the third paragraph.

%%%%%%%%
\point{Fig. 4:}{
    In the caption, I think it’s misleading to say that $\migrate$ is ``relatively insensitive to $s_m$''. 
    It’s exponentially sensitive! I don’t understand the blue numbers in the right-hand
    panel. It might be easier to interpret these if they were contour plots
    of the total time to adaptation, colored according to the value of (14),
    although I’m not sure. Also, they aren’t really ``phase diagrams''.
}

\reply
Good points; we meant that it was less sensitive to $s_m$ than if it didn't have the square root;
in this draft we've replaced this figure with a different one (so
disgarded this comment).

@@

\reply
Good point; fixed.

%%%%%%%%
\point{}{ 
 In (56) and following, it would be clearer to use different symbols for scaled
and unscaled distances.

}

\reply
We've put the scaling directly back into the equations here in this
new draft, so this isn't an issue.

@@

\point{3.}{
    Section "length of shared haplotype": This is very handwavy and
    (without simulation backup) I find it hard to believe that these results
    are accurate. Since it is also not central to the paper, the section
    should maybe just be dropped.
}

\reply
We have dropped this section.

%%%%%%
\point{[beginning comments in the pdf]}{
``Do you assume a Gaussian kernel? Or is this irrelevant? Which properties of the kernel are required? Only existence of a finite $\sigma^2$?''
}

\reply
We've made this more clear here 
\llname{rr:longtailed}.

@@

        \item in your arguments, you use the outflow of families from a ball with (large) radius $r_0$ [although I do not see how this figures in]

        \item over short ranges, results must depend on details of the dispersal kernel anyway. [Depending on the shape of the distribution - heavy tail - this even holds for large distances]. As mentioned above: spell out assumptions on the kernel.
    \end{itemize}

So: what is the minimal distance $|y|$ for the result to be valid? And: does this minimal distance eliminate the anomaly of an increasing $\migrate$ with $s_m$ for small $s_m?$ Alternatively: is there also a lower limit for $s_m$? I have noted that you do assume above that $s_m$ is large enough to avoid local fixation - but I'm not sure that this comes down to the same thing. 
\end{quote}
}

\reply
As discssed above, we've added this discussion to the Section ``\nameref{ss:assumptions}''.
A few more notes:
It's not clear to us why the second point requires the patch width to be wider than $1/s_m$, shouldn't it be $\sigma/\sqrt{s_m}$
(which is covered elsewhere)?
The appearance of $r_0$ is purely to make families well-defined, and don't think it needs to be discussed here.

### 2014-12-09 patchy-selection-paper.tex
small changes throughout response to reviews, and application section

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
Top panel: The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
see text. Bottom panel: The initial genetic length of the haplotype
shared between patches due to adaptation via migration as a function
of the distance between environmental patches. 
  \label{fig:mice_prob_parallel}
  }
\end{figure}

@@

\gc{DELETE THIS: 
Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
we should expect the haplotype to be rapidly whittled down within
patches after establishment (equation \ref{eqn:rec_rate_within}). We
estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
(M.~Nachman, personal communication),
and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
is only expected to be on the order of hundreds or thousands of bases
($10^{-3}$--$10^{-4}$cM).
This suggests that unless the migration has happened fairly recently
then it will be difficult to identify shared haplotypes
(how recently depends on the background level of LD in the population).}

>>>>

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
Top panel: The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
see text. Bottom panel: The initial genetic length of the haplotype
shared between patches due to adaptation via migration as a function
of the distance between environmental patches. As a rough rule of
thumb $1$cM is approximately $1$Mb in a number of mammalian species. \linelabel{rr:cMMb}
  \label{fig:mice_prob_parallel}
  }
\end{figure}

@@

%\gc{DELETE THIS: 
% Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
% we should expect the haplotype to be rapidly whittled down within
% patches after establishment (equation \ref{eqn:rec_rate_within}). We
% estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
% Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
% (M.~Nachman, personal communication),
% and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
% then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
% is only expected to be on the order of hundreds or thousands of bases
% ($10^{-3}$--$10^{-4}$cM).
% This suggests that unless the migration has happened fairly recently
% then it will be difficult to identify shared haplotypes
% (how recently depends on the background level of LD in the population).}
